Dancing on Water: The Choreography of Sulfur Dioxide Adsorption to Aqueous Surfaces

### 2011-10-18 EquilibratedOrientation.tex
working on review responses

The bisector tilt of the water molecules, $\theta$, concentrates around $\theta=$ 90\textdegrees within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems. Above the surface location at positive distances, $\theta \leq$ 90\textdegrees~indicates that the water hydrogens tend to point towards the gas-phase side of the surface. As the tilt nears $\theta=$ 90\textdegrees~the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a similar trend where waters at or just below the surface have values of $\phi$ near 90\textdegrees, and waters above the surface take on values of $\phi$ near 0\textdegrees. This jump in the angular distribution of $\phi$ indicates that waters at or below the surface lie mostly flat in the plane of the interface, and as they move above the surface towards the gas phase, they reorient with one OH bond pointing towards the water bulk, and one pointing out of the surface into the gas. This behavior is more strongly pronounced in the neat-water system where most of the surface waters are not interacting with an adsorbed layer of \suldiox~molecules.

@@

Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase. For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom plot of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta leq$ 90\textdegrees. This results from waters with bisectors pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk. The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly centered around $\theta)=$ 0\textdegrees (see Figure \ref{fig:water-angles}).

@@

In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta \lt$ 90 \textdegrees. The peak of the distribution occurs at $\theta)=$ 0\textdegrees. This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase. This same distribution occurs in the saturated system for depths below the surface location, $\lt 0$\angs. Beyond 4\angs~above the surface, both distributions become mostly isotropic, either because the \suldiox~does not venture into the gas in the neat-water system, or because of the nature of the adsorbed \suldiox~layer in the saturated system. Promixity to the water surface highly orients the \suldiox~bisector. % with the sulfur atom pointing in towards the water bulk. %Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 4\angs. %Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.

>>>>

The bisector tilt of the water molecules, $\theta$, concentrates around $\theta=$ 90\textdegree within the first few\angs~above and below the water surface location, becoming progressively isotropic further through the interfacial region and into the water bulk of both systems. Above the surface location at positive distances, $\theta \leq$ 90\textdegree~indicates that the water hydrogens tend to point towards the gas-phase side of the surface. As the tilt nears $\theta=$ 90\textdegree~the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a similar trend where waters at or just below the surface have values of $\phi$ near 90\textdegree, and waters above the surface take on values of $\phi$ near 0\textdegree. This jump in the angular distribution of $\phi$ indicates that waters at or below the surface lie mostly flat in the plane of the interface, and as they move above the surface towards the gas phase, they reorient with one OH bond pointing towards the water bulk, and one pointing out of the surface into the gas. This behavior is more strongly pronounced in the neat-water system where most of the surface waters are not interacting with an adsorbed layer of \suldiox~molecules.

@@

Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the orientation of those waters furthest into the gas phase. For the saturated solution, the resulting orientation of waters above 0\angs, shown in the bottom plot of Figure \ref{fig:water-orientation}, is nearly isotropic in $\phi$, and with $\theta leq$ 90\textdegree. This results from waters with bisectors pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk. The effect is more pronounced as the waters move further from the water surface, and above 4\angs~the $\theta$ distribution is mostly centered around $\theta)=$ 0\textdegree (see Figure \ref{fig:water-angles}).

@@

In the interfacial region the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\theta \lt$ 90 \textdegree. The peak of the distribution occurs at $\theta)=$ 0\textdegree. This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase. This same distribution occurs in the saturated system for depths below the surface location, $\lt 0$\angs. Beyond 4\angs~above the surface, both distributions become mostly isotropic, either because the \suldiox~does not venture into the gas in the neat-water system, or because of the nature of the adsorbed \suldiox~layer in the saturated system. Promixity to the water surface highly orients the \suldiox~bisector. % with the sulfur atom pointing in towards the water bulk. %Moving further away from the water surface, and interacting less with \wat~molecules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 4\angs. %Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.

### 2011-10-18 Methods.tex
working on review responses

The low and high concentration simulated systems, ``neat-water'' and ``saturated water'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration. Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface. The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure. The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

>>>>

The low and high concentration simulated systems, ``neat-water'' and ``saturated water'', respectively, were created as follows: the neat-water simulation involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~neat-water system was then evolved for 2 ns to produce an equilibrated starting configuration. Because of the extremely low concentration of \suldiox~in the neat-water system the surface waters behave similarly to a true neat-water air-liquid interface. The neat-water system orientational results shown later in this work reproduce well the results of our previous orientational studies of surface water behavior.\cite{Walker2006b,Hore2008} The saturated system had 22 \suldiox~molecules introduced to the water slab bulk in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water (k\textdegree$_H = 1.4$ mol/kg*bar).\cite{Lide2000} Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the saturated water slab to simulate an added \suldiox~gas pressure. The saturated system with both bulk and gaseous \suldiox~was then evolved for 2 ns to produce a starting configuration for further saturated simulations.
